---
title: "Exploring fitLHS results"
output: html_document
---

```{r load, echo = FALSE}
fn <- "fitLHS.rda"
fn2 <- "fitLHS2.rda"
load(fn)
load(fn2)
```

I performed 500 simulations based on these parameters:

```{r pars}
colnames(p) <- c("log.beta", "log.gamma", "log.N", "logit.i")
pairs(~log.beta+log.gamma+log.N+logit.i, data = as.data.frame(p))

```

Does `fitsir.optim` perform better than `fitsir`? I put the parameters in an increasing order to see if we get a stair-like shape:

```{r performe1,echo = FALSE}
par(mfrow = c(1,2))
matplot(apply(fpars, 2, sort), type = "l", ylab = "param value", main = "fitsir")
text(8,tail(fpars,1),colnames(fpars),col=1:4,adj=-1)
matplot(apply(spars, 2, sort), type = "l", ylab = "param value", main = "fitsir.optim")
text(8,tail(fpars,1),colnames(fpars),col=1:4,adj=-1)
```

Can we look at the histograms for negative log likelihood for both sims?

```{r performace2, echo = FALSE}
par(mfrow = c(1,2))
hist(f.nll, main = "fitsir")
hist(s.nll, main = "fitsir.optim")
```

It seems like `fitsir` performs better, although only half of them gives us good fits... Also, we do get a lot of errors with `fitsir.optim`:

```{r error}
sum(is.na(spars[,1]))
```

Let's just look at `fitsir` reults for now... I actually tried taking `fpars` parameters as a starting point and try running `fitsir` with them. Let's see how performance has improved:

```{r hist, echo = FALSE}
par(mfrow = c(1,2))
hist(f.nll)
hist(f.nll2)
```

We can also look at how negative log likelihood values have changed after the second fit.

```{r fit.compare}
nll.df <- data.frame(nll1 = f.nll, nll2 = f.nll2)
nll.df <- nll.df[order(nll.df[,1]),]
par(mfrow = c(1,1))
plot(nll.df)
abline(a=0,b=1)
```

Does this suggest that some starting values cannot be optimized? Let's look into the preliminary results more... I'm going to sort out the good fits:

```{r goodfits}
fpars.good <- fpars[(f.nll < 169),]
f.nll.good <- f.nll[(f.nll < 169)]

```

This is quite interesting...

```{r}
pairs(~log.beta+log.gamma+log.N+logit.i, data = as.data.frame(fpars.good))
```


```{r eval = FALSE, echo = FALSE}

##terrible fit
findSens(bombay2, fpars[5,], plot.it = TRUE)
findSens(bombay2, spars[5,], plot.it = TRUE)
##running it once more brings it to a local minimum
tmp.pars <- fitsir.optim(bombay2, start = spars[5,], plot.it = TRUE)
tmp.pars2 <- fitsir(bombay2, start = fpars[5,])

```