---
title: "Exploring fitLHS results"
output: html_document
---

```{r pkgs,echo=FALSE,message=FALSE}
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2); theme_set(theme_bw())
```
```{r load, echo = FALSE}
fn <- "fitLHS.rda"
fn2 <- "fitLHS2.rda"
## BMB: what's the difference?
## SWP: L2 uses L1 as starting values to fit the curve
L1 <- load(fn)
L2 <- load(fn2)
```

I performed 500 simulations based on these parameters:

```{r pars}
colnames(p) <- c("log.beta", "log.gamma", "log.N", "logit.i")
pairs(~log.beta+log.gamma+log.N+logit.i, data = as.data.frame(p),
      gap=0,cex=0.5)
```

Does `fitsir.optim` perform better than `fitsir`? I put the parameters in an increasing order to see if we get a stair-like shape:

```{r plotranks,echo=FALSE}
dd <- rbind(data.frame(method="fitsir",NLL1=f.nll,NLL2=f.nll2),
      data.frame(method="fitsir.optim",NLL1=s.nll,NLL2=s.nll2)) %>%
    gather(run,NLL,-method) %>%
    group_by(method,run) %>%
    mutate(r=rank(NLL))
ggplot(dd,aes(r,NLL,colour=method,linetype=run))+geom_line()+
    labs(x="rank",y="negative log-likelihood")+
    scale_colour_brewer(palette="Set1")
```

How many NA values?

```{r,natab}
with(dd,table(is.na(NLL),run,method))
```

```{r performe1,echo = FALSE}
par(mfrow = c(1,2))
matplot(apply(fpars, 2, sort), type = "l", ylab = "param value", main = "fitsir")
text(8,tail(fpars,1),colnames(fpars),col=1:4,adj=-1)
matplot(apply(spars, 2, sort), type = "l", ylab = "param value", main = "fitsir.optim")
text(8,tail(fpars,1),colnames(fpars),col=1:4,adj=-1)
```

Can we look at the histograms for negative log likelihood for both sims?

```{r performace2, echo = FALSE}
par(mfrow = c(1,2))
hist(f.nll, main = "fitsir")
hist(s.nll, main = "fitsir.optim")
```

It seems like `fitsir` performs better, although only half of them give us good fits... Let's just look at `fitsir` results for now... I actually tried taking `fpars` parameters as a starting point and try running `fitsir` with them. Let's see how performance has improved:

```{r hist, echo = FALSE}
par(mfrow = c(1,2))
hist(f.nll)
hist(f.nll2)
```

We can also look at how negative log likelihood values have changed after the second fit.

```{r fit.compare}
nll.df <- data.frame(nll1 = f.nll, nll2 = f.nll2)
nll.df <- nll.df[order(nll.df[,1]),]
par(mfrow = c(1,1))
plot(nll.df)
abline(a=0,b=1)
```

Does this suggest that some starting values cannot be optimized? Let's look into the preliminary results more...

Let's try Raue et al graph:

```{r raue}
tmpfun <- function(n.sample, i1, i2){
    set.seed(101)
    s.n <- sample(c(1:500), n.sample)
    p.sample <- p[s.n,]
    fpars.sample <- fpars[s.n,]
    plot(p.sample[,c(i1,i2)])
    points(fpars.sample[,c(i1,i2)], col = 2)
    arrows(p.sample[,i1], p.sample[,i2], x1=fpars.sample[,i1], y1 = fpars.sample[,i2], col = "Gray")
}

par(mfrow = c(2,3))
tmpfun(30, 1, 2)
tmpfun(30, 1, 3)
tmpfun(30, 1, 4)
tmpfun(30, 2, 3) 
tmpfun(30, 2, 4)
tmpfun(30, 3, 4) 
```

I'm going to sort out the good fits:

```{r goodfits}
fpars.good <- fpars[(f.nll < 169),]
f.nll.good <- f.nll[(f.nll < 169)]
```

This is quite interesting...

```{r}
pairs(~log.beta+log.gamma+log.N+logit.i, data = as.data.frame(fpars.good),
      gap=0,cex=0.5)
```

Let's try plotting all of them:

```{r good.fits2, error = FALSE, message=FALSE}
library(fitsir)
par(mfrow = c(1,1))
hist(f.nll.good)

bombay2 <- setNames(bombay, c("tvec", "count"))
tmpfun2 <- function(i, pars){
    I <- SIR.detsim(bombay2$tvec, trans.pars(pars[i,]))
}
library(reshape2)
tmpfun3 <- function(n, pars){
    good.sims <- list()
    good.sims <- lapply(c(1:n),function(i){data.frame(tvec = bombay2$tvec, count = tmpfun2(i, pars))})
    mL <- melt(good.sims, id.vars = "tvec")
ggplot(bombay2, aes(tvec, count)) + geom_point() +
    geom_line(data = mL, aes(tvec, value, group = L1), col = "Gray")
}

n <- nrow(fpars.good)
tmpfun3(n, fpars.good)
```

```{r eval = FALSE, echo = FALSE}
##terrible fit
findSens(bombay2, fpars[5,], plot.it = TRUE)
findSens(bombay2, spars[5,], plot.it = TRUE)
##running it once more brings it to a local minimum
tmp.pars <- fitsir.optim(bombay2, start = spars[5,], plot.it = TRUE)
tmp.pars2 <- fitsir(bombay2, start = fpars[5,])
```

