---
title: "Sensitivity demo"
output: html_document
---

```{r pkgs, message = FALSE}
library("deSolve")
library("bbmle")
library("devtools")
load_all("..")
```

I'm going to show how to use the sensitivity functions...

## Gradient function

I'm going to claim that this is the correct sensitivity equation:

$$
\begin{aligned}
\frac{\partial}{\partial t}\frac{\partial I(t,\theta)}{\partial\theta} &= \frac{\partial f_I}{\partial\theta}\\
&= \frac{\partial f_I}{\partial S}\frac{\partial  S}{\partial\theta} + \frac{\partial f_I}{\partial I}\frac{\partial I}{\partial\theta} + \frac{\partial f_I}{\partial\theta} \\
\end{aligned}
$$

## Example 

```{r ex.sim}
tvec <- seq(1, 60, 0.2)

pars <- c(log.beta = -0.7,
	log.gamma = -2.3,
	log.N= 6.21,
	logit.i = -4.6)

pars2 <- c(log.beta = -0.7,
	log.gamma = -2.3,
	log.N= 6.21,
	logit.i = -4.5999)

a = SIR.detsim(tvec, trans.pars(pars), findSens = TRUE)

a2 = SIR.detsim(tvec, trans.pars(pars2), findSens = TRUE)

z = (exp(a2$logI)-exp(a$logI))/0.0001

plot(exp(a$logI), type = "l")
lines(exp(a2$logI), col = "2")

plot(z, type = "l")

lines(a$nu_I0_I, col = 2)

lines(a2$nu_I0_I, col = 3)

```

## Bombay example - initial values

We start with these three equations:

$$
\begin{aligned}
r &= \beta - \gamma,\\
-\frac{Q}{I_p} &= \frac{\beta\gamma}{N},\\
N - I_0 - \int I(t) dt = \frac{\gamma}{\beta}N,
\end{aligned}
$$

where $I(t)$ is incidence. Substition gives us following two equations:

$$
\begin{aligned}
N &= - \gamma^2 \frac{I_p}{Q} + I_0 + \int I(t) dt\\
N &= - (\gamma+r) \gamma \frac{I_p}{Q}
\end{aligned}
$$

This gives

$$
\begin{aligned}
(\gamma+r) \gamma \frac{I_p}{Q} &= \gamma^2 \frac{I_p}{Q} - I_0 - \int I(t) dt\\
\gamma^2 \frac{I_p}{Q} + r \gamma \frac{I_p}{Q} &= \gamma^2 \frac{I_p}{Q} - I_0 - \int I(t) dt\\
r \gamma \frac{I_p}{Q} &=- (I_0 + \int I(t) dt)\\
\end{aligned}
$$

```{r}
library("fitsir")

bombay2 <- setNames(bombay,c("tvec","count"))

iniP.ex <- startfun(data=bombay2, auto = TRUE)

findSens(bombay2, iniP.ex, plot.it = TRUE)

(fitP <- coef(fitsir(bombay2, start = startfun(data = bombay2, auto = TRUE))))

findSens(bombay2, fitP, plot.it = TRUE)

```

```{r, error = FALSE}
(fitP2 <- fitsir.optim(bombay2, start = startfun(), plot.it=TRUE))

findSens(bombay2, fitP2, plot.it = TRUE)

##with auto start?

(fitP3 <- fitsir.optim(bombay2, start = startfun(data=bombay2, auto = TRUE), plot.it=TRUE))

findSens(bombay2, fitP3, plot.it = TRUE)

```

```{r }

(fitP4 <- fitsir.optim(bombay2, start = fitP3, plot.it=TRUE))

findSens(bombay2, fitP4, plot.it = TRUE)

```

## Incidence

We need new equations for incidence... Let's assume that bombay2 data is incidence data. We're going to use $I(t)$ for incidence and $P(t)$ for prevalence from here.

```{r}

bombay2 <- setNames(bombay,c("tvec","count"))

fitP.inc <- coef(fitsir(bombay2, incidence = TRUE))

findSens(bombay2, fitP.inc, plot.it = TRUE, incidence = TRUE)

bombay3 <- data.frame(tvec = bombay2$tvec, count = cumsum(bombay2$count))

plot(bombay3)

lines(cumsum(SIR.detsim(bombay2$tvec, trans.pars(fitP.inc), incidence = TRUE)), col = 2)
```

### Sensitivity

```{r inc.sens}
tvec <- seq(1, 50, 0.5)

trans.pars(pars)

fake.inc = data.frame(tvec, count = SIR.detsim(tvec, trans.pars(pars), incidence = TRUE))

plot(fake.inc)

fake.P <- coef(fitsir(fake.inc, incidence = TRUE))

trans.pars(fake.P)

findSens(fake.inc, fake.P, plot.it = TRUE, incidence = TRUE)
```

```{r}
(fake.P2 <- fitsir.optim(fake.inc, start = startfun(), plot.it=TRUE, incidence = TRUE))

findSens(fake.inc, fake.P2, plot.it = TRUE, incidence = TRUE)

```
