---
title: "Sensitivity demo"
output: html_document
---

```{r pkgs, message = FALSE}
library("deSolve")
library("bbmle")
source("../R/fitSIR_funs.R")
```

I'm going to show how to use the sensitivity functions...

## Gradient function

I'm going to claim that this is the correct sensitivity equation:

$$
\begin{aligned}
\frac{\partial}{\partial t}\frac{\partial I(t,\theta)}{\partial\theta} &= \frac{\partial f_I}{\partial\theta}\\
&= \frac{\partial f_I}{\partial S}\frac{\partial  S}{\partial\theta} + \frac{\partial f_I}{\partial I}\frac{\partial I}{\partial\theta} + \frac{\partial f_I}{\partial\theta} \\
\end{aligned}
$$

## Example 

```{r ex.sim}
tvec <- seq(1, 60, 0.2)

pars <- c(beta = 0.5,
	gamma = 0.1,
	N= 501,
	i0 = 0.01)

pars2 <- c(beta = 0.505,
	gamma = 0.1,
	N= 501,
	i0 = 0.01)

a = SIR.detsim(tvec, pars, findSens = TRUE)

a2 = SIR.detsim(tvec, pars2, findSens = TRUE)

z = (a2$I-a$I)/0.005

plot(a$I, type = "l")
lines(a2$I, col = "2")

plot(z, type = "l", ylim = c(-500, 1000))

lines(a$nu_beta_I, col = 2)

lines(a2$nu_beta_I, col = 3)

```

## SSQ example

```{r ssq.ex}
fake.obs = as.data.frame(cbind(tvec, I=a[,c("I")]))
colnames(fake.obs) = c("tvec", "count")

pars3 <- c(beta = 0.51,
	gamma = 0.10,
	N= 501,
	i0 = 0.01)

pars4 <- c(beta = 0.501,
	gamma = 0.10,
	N= 501,
	i0 = 0.01)

findSens(fake.obs, pars3)
findSens(fake.obs, pars4)

```

## Bombay example - initial values

We start with these three equations:

$$
\begin{aligned}
r &= \beta - \gamma,\\
-\frac{Q}{I_p} &= \frac{\beta\gamma}{N},\\
N - I_0 - \int I(t) dt = \frac{\gamma}{\beta}N,
\end{aligned}
$$

where $I(t)$ is incidence. Substition gives us following two equations:

$$
\begin{aligned}
N &= - \gamma^2 \frac{I_p}{Q} + I_0 + \int I(t) dt\\
N &= - (\gamma+r) \gamma \frac{I_p}{Q}
\end{aligned}
$$

This gives

$$
\begin{aligned}
(\gamma+r) \gamma \frac{I_p}{Q} &= \gamma^2 \frac{I_p}{Q} - I_0 - \int I(t) dt\\
\gamma^2 \frac{I_p}{Q} + r \gamma \frac{I_p}{Q} &= \gamma^2 \frac{I_p}{Q} - I_0 - \int I(t) dt\\
r \gamma \frac{I_p}{Q} &=- (I_0 + \int I(t) dt)\\
\end{aligned}
$$

```{r}
library("fitsir")

bombay2 <- setNames(bombay,c("tvec","count"))

iniP.ex <- trans.pars(startfun(data=bombay2, auto = TRUE))

findSens(bombay2, iniP.ex, plot.it = TRUE)

fitP <- coef(fitsir(bombay2, start = startfun(data=bombay2, auto = TRUE)))

(fit.pars <- trans.pars(fitP))

findSens(bombay2, fit.pars, plot.it = TRUE)

```

```{r, error = FALSE}
(fitP2 <- fitsir.optim(bombay2, start = startfun(), plot.it=TRUE))

findSens(bombay2, fitP2, plot.it = TRUE)

##with auto start?

(fitP3 <- fitsir.optim(bombay2, start = startfun(data=bombay2, auto = TRUE), plot.it=TRUE))

findSens(bombay2, fitP3, plot.it = TRUE)

```

This is weird... We get lower sensitivity values but higher SSQ...

```{r }

(fitP4 <- fitsir.optim(bombay2, start = fitP, plot.it=TRUE))

findSens(bombay2, fitP4, plot.it = TRUE)

```

## Incidence

We need new equations for incidence... Let's assume that bombay2 data is incidence data. We're going to use $I(t)$ for incidence and $P(t)$ for prevalence from here.

```{r}

bombay2 <- setNames(bombay,c("tvec","count"))

fitP.inc <- coef(fitsir(bombay2, incidence = TRUE))

(fit.pars2 <- trans.pars(fitP.inc))

findSens(bombay2, fit.pars2, plot.it = TRUE, incidence = TRUE)

bombay3 <- data.frame(tvec = bombay2$tvec, count = cumsum(bombay2$count))

plot(bombay3)

lines(cumsum(SIR.detsim(bombay2$tvec, fit.pars2, incidence = TRUE)), col = 2)
```

I think I made incidence fitting to work but I still want to figure out the starting parameters.

```{r}

plot(SIR.detsim(1:800, fit.pars2, incidence = TRUE))

```

### Auto-start

First, we know that $I(0)/ \Delta t = \beta N (1-i_0) i_0 $, where $\Delta t$ is the reporting time interval.
```{r}
tvec <- bombay2$tvec
count <- bombay2$count

count[1] #4
with(as.list(fit.pars),{beta * N * (1 -i0) * i0}) ## 4.35
```

We also know that the the initial exponential rate of growth for $\int I(t) dt$ is $\beta$.

```{r}

ss <- smooth.spline(tvec,log(count),spar=0.5)
		## find max value
ss.tmax <- uniroot(function(x) predict(ss,x,deriv=1)$y,range(tvec))$root
		## find a point halfway between initial and max
		##  scaling could be adjustable?
ss.thalf <- min(tvec)+0.5*(ss.tmax-min(tvec))

m1 <- lm(log(count)~tvec, data =bombay3[ss.thalf:ss.tmax,])

coef(m1)[2] ## 0.28
fit.pars[1] ## 0.372
```

### Sensitivity

```{r inc.sens}
tvec <- seq(1, 60, 0.2)
pars <- c(beta = 0.5,
	gamma = 0.1,
	N= 501,
	i0 = 0.01)

pars2 <- c(beta = 0.5,
	gamma = 0.101,
	N= 501,
	i0 = 0.01)

a = SIR.detsim(tvec, pars, findSens = TRUE, inc = TRUE)

a2 = SIR.detsim(tvec, pars2, findSens = TRUE, inc = TRUE)

z = (a2$inc-a$inc)/0.001

plot(a$inc, type = "l")
lines(a2$inc, col = "2")

plot(z, type = "l", ylim = c(-50, 50))

lines(a$nu_gamma_I, col = 2)

lines(a2$nu_gamma_I, col = 3)
```

