---
title: "Hessian"
output: html_document
---

I want to get used to using `numDeriv` first... Here's an arbitrary function:

```{r}
f <- function(par){
    with(as.list(par),{
        x^2 * y + 2 * x * y^3 - x 
    })
}

```

It's gradient function:

```{r}
f.g <- function(par){
    with(as.list(par),{
        dfx <- 2 * x * y + 2 * y^3 - 1
        dfy <- x^2 + 6 * x * y^2
        return(c(dfx, dfy))
    })
}

```

Example:

```{r}

library("numDeriv")
pars <- c(x = 2, y =1)

jacobian(f, pars)
f.g(pars)

```

Hessian function:

```{r}

f.g2 <- function(par){
    with(as.list(pars),{
        dfxx <- 2 * y
        dfxy <- 2 * x + 6 * y^2
        dfyx <- 2 * x + 6 * y^2
        dfyy <- 12 * x * y
        return(c(dfxx, dfxy, dfyx, dfyy))
    })
}

```

Does this work well?

```{r}

hessian(f, pars)
jacobian(f.g, pars)
f.g2(pars)

```

I think I know how it works now...

## SIR

```{r load, error = FALSE, message = FALSE}
library(bbmle)
library(fitsir)
library("deSolve")
source("../R/fitSIR_funs.R")

bombay2 <- setNames(bombay, c("tvec", "count"))

fpars <- coef(fitsir(bombay2, start = startfun()))

jacobian(findSSQ, fpars, data = bombay2, SSQonly = TRUE)

findSens(bombay2, fpars, sensOnly = TRUE)

```

Jacobian seems to work fairly well. Can we find the Hessian?

```{r}
(hess.m <- hessian(findSSQ, fpars, data = bombay2, SSQonly = TRUE))
(hess.grad <- jacobian(findSens, fpars, data = bombay2, sensOnly = TRUE))

eigen(hess.m)
eigen(hess.grad)
```

OK... These should give us the same values but they're not... I'm assuming that `hess.grad` should be more accurate so if that is correct, we do have a local mimimum... It would be still nice to know if we can calculate the hessian algebraically... Here's the model:

$$
\begin{aligned}
&\frac{dSSQ}{d\log \theta}\\
&= \frac{dSSQ}{d\theta} \frac{d\theta}{d\log \theta}\\
&= \frac{dSSQ}{dI} \frac{dI}{d\theta} \frac{d\theta}{d\log\theta}\\
&= 2(I-\hat{I}) \frac{dI}{d\theta} \theta
\end{aligned}
$$

Second derivative:

$$
\begin{aligned}
&\frac{d(dSSQ/d\log\theta_1)}{d\log\theta_2}\\
&= \frac{d(dSSQ/d\log\theta_1)}{d\theta_2} \theta_2\\
&= (2\frac{dI}{d\theta_2} \frac{dI}{d\theta_1} \theta_1 + 2(I-\hat{I}) \frac{d^2I}{d\theta_1\theta_2}\theta_1 + 2(I-\hat{I})\frac{dI}{d\theta_1} \frac{d\theta_1}{d\theta_2} ) \theta_2\\
\end{aligned}
$$

```{r}
SIR.grad.hessian <- function(t, x, params) {
    g <- with(as.list(c(x,params)),
    {
        I <- exp(logI)
        dS <- -beta*S*I/N
        dlogI <- beta*S/N-gamma
        
        grad_SS <- - beta * I/N
        grad_SI <- - beta * S/N
        grad_IS <- beta*I/N
        grad_II <- beta*S/N-gamma
    
        ##Sens for beta 
        dnu_S_b <- grad_SS * nu_S_b + grad_SI * nu_I_b - S*I/N
        ##Need 4 more equations
        dnu_S_bb <- (- I/N - beta/N * nu_I_b) * nu_S_b + grad_SS * nu_S_bb +  (- S/N - beta/N * nu_S_b) * nu_I_b + grad_SI * nu_I_bb - 
            I/N * nu_S_b - S/N * nu_I_b
        
        dnu_S_bg <- (-beta/N * nu_I_g) * nu_S_b + grad_SS * nu_S_bg + (-beta/N * nu_S_g) * nu_I_b + grad_SI * nu_I_bg -
            I/N * nu_S_g - S/N * nu_I_g
        
        dnu_S_bN <- (beta*I/N^2 - beta/N * nu_I_N) * nu_S_b + grad_SS * nu_S_bN + (beta*S/N^2 - beta/N * nu_S_N) * nu_I_b + grad_SI * nu_I_bN +
            S * I/N^2 - I/N * nu_S_N - S/N * nu_I_N
        
        dnu_S_bi <- (-beta/N * nu_I_i) * nu_S_b + grad_SS * nu_S_bi + (-beta/N * nu_S_i) * nu_I_b + grad_SI * nu_I_bi -
            I/N * nu_S_i - S/N * nu_I_i
        
        ##Sens for gamma
        dnu_S_g <- grad_SS * nu_S_g + grad_SI * nu_I_g
        ##Need 3 more equations
        dnu_S_gg <- (-beta/N * nu_I_g) * nu_S_g + grad_SS * nu_S_gg + (-beta/N * nu_S_g) * nu_I_g + grad_SI * nu_I_gg
        
        dnu_S_gN <- (beta*I/N^2 - beta/N * nu_I_N) * nu_S_g + grad_SS * nu_S_gN + (beta*S/N^2 - beta/N * nu_S_N) * nu_I_g + grad_SI * nu_I_gN
        
        dnu_S_gi <- (-beta/N * nu_I_i) * nu_S_g + grad_SS * nu_S_gi + (-beta/N * nu_S_i) * nu_I_g + grad_SI * nu_I_gi
        
        ##Sens for N
        dnu_S_N <- grad_SS * nu_S_N + grad_SI * nu_I_N + beta*S*I/N^2
        ##Need 2 more equations
        ##dnu_S_NN <- (beta*I/N^2 - beta/N * nu_I_N) * nu_S_N + grad_SS * nu_S_NN + (beta*S/N^2 - beta/N * nu_S_N) * nu_I_N + grad_SI * nu_I_NN -
        ##    2*beta*S*I/N^3 + beta*S/N^2 * nu_I_N + beta*I/N^2 * nu_S_N
        dnu_S_NN <- 0
        
        dnu_S_Ni <- (-beta/N * nu_I_i) * nu_S_N + grad_SS * nu_S_Ni + (-beta/N * nu_S_i) * nu_I_N + grad_SI * nu_I_Ni +
            beta*S/N^2 * nu_I_i + beta*I/N^2 * nu_S_i
        
        ##Sens for i0
        dnu_S_i <- grad_SS * nu_S_i + grad_SI * nu_I_i
        ##Need 1 more equaion
        dnu_S_ii <- (-beta/N * nu_I_i) * nu_S_i + grad_SS * nu_S_ii + (-beta/N * nu_S_i) * nu_I_i + grad_SI * nu_I_ii
        
        ##Sens for beta
        dnu_I_b <- grad_IS * nu_S_b + grad_II * nu_I_b + S*I/N
        ##Need 4 more equations
        dnu_I_bb <- (I/N + beta/N * nu_I_b) * nu_S_b + grad_IS * nu_S_bb + (S/N + beta/N * nu_S_b) * nu_I_b + grad_II * nu_I_bb +
            I/N * nu_S_b + S/N * nu_I_b
        
        dnu_I_bg <- (beta/N * nu_I_g) * nu_S_b + grad_IS * nu_S_bg + (-1 + beta/N * nu_S_g) * nu_I_b + grad_II * nu_I_bg + 
            I/N * nu_S_g + S/N * nu_I_g
        
        dnu_I_bN <- (-beta*I/N^2 + beta/N * nu_I_N) * nu_S_b + grad_IS * nu_S_bN + (-beta*S/N^2 + beta/N * nu_S_N) * nu_I_b + grad_II * nu_I_bN -
            S * I/N^2 + I/N * nu_S_N + S/N * nu_I_N
        
        dnu_I_bi <- (beta/N * nu_I_i) * nu_S_b + grad_IS * nu_S_bi + (beta/N * nu_S_i) * nu_I_b + grad_II * nu_I_bi +
            I/N * nu_S_i + S/N * nu_I_i
        
        ##Sens for gamma
        dnu_I_g <- grad_IS * nu_S_g + grad_II * nu_I_g - I
        ##Need 3 more equations
        dnu_I_gg <- (beta/N * nu_I_g) * nu_S_g + grad_IS * nu_S_gg + (-1 + beta/N * nu_S_g) * nu_I_g + grad_II * nu_I_gg -
            nu_I_g
        
        dnu_I_gN <- (-beta*I/N^2 + beta/N * nu_I_N) * nu_S_g + grad_IS * nu_S_gN + (-beta*S/N^2 + beta/N * nu_S_N) * nu_I_g + grad_II * nu_I_gN -
            nu_I_N
        
        dnu_I_gi <- (beta/N * nu_I_i) * nu_S_g + grad_IS * nu_S_gi + (beta/N * nu_S_i) * nu_I_g + grad_II * nu_I_gi -
            nu_I_i
        
        ##Sens for N
        dnu_I_N <- grad_IS * nu_S_N + grad_II * nu_I_N - beta*S*I/N^2
        ##Need 2 more equations
        ##dnu_I_NN <- (-beta*I/N^2 + beta/N * nu_I_N) * nu_S_N + grad_IS * nu_S_NN + (-beta*S/N^2 + beta/N * nu_S_N) * nu_I_N + grad_II * nu_I_NN +
        ##    2*beta*S*I/N^3 - beta*S/N^2 * nu_I_N - beta*I/N^2 * nu_S_N
        dnu_I_NN <- 0
        
        dnu_I_Ni <- (beta/N * nu_I_i) * nu_S_N + grad_IS * nu_S_Ni + (beta/N * nu_S_i) * nu_I_N + grad_II * nu_I_Ni -
            beta*S/N^2 * nu_I_i - beta*I/N^2 * nu_S_i
        
        ##Sens for i0
        dnu_I_i <- grad_IS * nu_S_i + grad_II * nu_I_i
        ##Need 1 more equaion
        dnu_I_ii <- (beta/N * nu_I_i) * nu_S_i + grad_IS * nu_S_ii + (beta/N * nu_S_i) * nu_I_i + grad_II * nu_I_ii
           
	
        list(c(dS, dlogI,
               dnu_S_b, dnu_S_bb, dnu_S_bg, dnu_S_bN, dnu_S_bi,
               dnu_S_g, dnu_S_gg, dnu_S_gN, dnu_S_gi,
               dnu_S_N, dnu_S_NN, dnu_S_Ni,
               dnu_S_i, dnu_S_ii,
               dnu_I_b, dnu_I_bb, dnu_I_bg, dnu_I_bN, dnu_I_bi,
               dnu_I_g, dnu_I_gg, dnu_I_gN, dnu_I_gi,
               dnu_I_N, dnu_I_NN, dnu_I_Ni,
               dnu_I_i, dnu_I_ii), I = I)
    })
}

SIR.detsim.hessian <- function(t, params){
    with(as.list(c(params)),{
        yini <- c(S = N*(1-i0), logI = log(N*i0),
                  nu_S_b = 0, nu_S_bb = 0, nu_S_bg = 0, nu_S_bN = 0, nu_S_bi = 0,
                  nu_S_g = 0, nu_S_gg = 0, nu_S_gN = 0, nu_S_gi = 0,
                  nu_S_N = 1-i0, nu_S_NN = 0, nu_S_Ni = -1,
                  nu_S_i = -N, nu_S_ii = 0,
                  nu_I_b = 0, nu_I_bb = 0, nu_I_bg = 0, nu_I_bN = 0, nu_I_bi = 0,
                  nu_I_g = 0, nu_I_gg = 0, nu_I_gN = 0, nu_I_gi = 0,
                  nu_I_N = i0, nu_I_NN = 0, nu_I_Ni = 1,
                  nu_I_i = N, nu_I_ii = 0
                  )
        odesol <- as.data.frame(rk(y=yini,
                                    times=t,
                                    func=SIR.grad.hessian,
                                    parms=params))
    })
    
}

tvec <- seq(1,30,0.1)

pars <- c(
    beta = 0.5,
    gamma = 0.1,
    N = 2,
    i0 = 0.01
)

pars2 <- c(
    beta = 0.5,
    gamma = 0.1,
    N = 3,
    i0 = 0.01
)

r1 <- SIR.detsim.hessian(tvec, pars)

r2 <- SIR.detsim.hessian(tvec, pars2)

matplot(cbind(r1$S, r2$S), type = "l")

plot((r2$S-r1$S)/1, type = "l")
lines(r1$nu_S_N, col = 2)
lines(r2$nu_S_N, col = 3)

nu_I_bg = (r2$nu_S_N-r1$nu_S_N)/1

plot(nu_I_bg, type = "l")
lines(r1$nu_S_NN, col = 2)
lines(r2$nu_S_NN, col = 3)

##Things that work

##b, bb, bg, bN, bi
##g, gg, gN, gi
##N, __ ,Ni 
##i, ii

##Not sure if NN is not working or if it's just a numerical error...

```

True hessian function:

```{r}

findHess <- function(data, params){
    t <- data$tvec
    tpars <- trans.pars(params)
    r <- SIR.detsim.hessian(t, tpars)
    
    with(as.list(c(tpars, r)),{
        
        parVec <- c(beta, gamma, N, i0)
        
        sensVec <- c(beta, gamma, N, i0^2*exp(-qlogis(i0)))
        
        sensVec2 <- c(1, 1, 1, 2*i0*exp(-qlogis(i0)) - i0^2 * exp(-qlogis(i0)) * 1/(i0-i0^2))
        
        derVec <- c("nu_I_b", "nu_I_g", "nu_I_N", "nu_I_i")
        
        derMat <- matrix(c("nu_I_bb", "nu_I_bg", "nu_I_bN", "nu_I_bi",
                           "nu_I_bg", "nu_I_gg", "nu_I_gN", "nu_I_gi",
                           "nu_I_bN", "nu_I_gN", "nu_I_NN", "nu_I_Ni",
                           "nu_I_bi", "nu_I_gi", "nu_I_Ni", "nu_I_ii"), 4, 4)
        
        findDeriv <- function(n1, n2){
            d1 <- get(derVec[n1])
            d2 <- get(derVec[n2])
            db <- get(derMat[n1,n2])
            
            if(n1 == n2){
                d12 <- sensVec2[n1]
            }else{
                d12 <- 0
            }
            deriv <- 2 * sum(d1 * d2 * sensVec[n1] + (I-data$count) * db * sensVec[n1] + (I-data$count) * d1 * d12) * sensVec[n2]
            return(deriv)
        }
        
        m <- matrix(NA, 4, 4)
        
        for(i in 1:4){
            for(j in 1:4){
                m[i,j] = findDeriv(i,j)
            }
        }
        
        return(m)
    })
    
}

(hess.alg <- findHess(bombay2, fpars))

(hess.grad <- jacobian(findSens, fpars, data = bombay2, sensOnly = TRUE))

all.equal(hess.alg, hess.grad)
```

I got the hessians to work.

```{r}

eigen(hess.alg)
##Positive definite.

fpars2 <- fitsir.optim(bombay2, start = startfun())

(hess.alg2 <- findHess(bombay2, fpars2))
(hess.m2 <- hessian(findSSQ, fpars2, data = bombay2, SSQonly = TRUE))
(hess.grad2 <- jacobian(findSens, fpars2, data = bombay2, sensOnly = TRUE))

eigen(hess.alg2)
##Another local minima?

```