---
title: "Stochastic sims"
author: Ben Bolker
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pkgs, message=FALSE}
library(fitsir)
library(bbmle) ## needed for coef() ...
library(splines) ## for ns()
library(plyr)  ## for raply()
```

```{r stochsimfun}
##' Stochastic simulation (observation/uncorrelated error only)
##' 
##' should this be called SIR.stochsim instead?
##' allow facility for (discrete and/or continuous time) process-error
##'   sims?
##'
##' @param pars parameters (on "original" constrained scale)
##' @param tmax max times
##' @param dt time step
##' @param rfun function for generating random (observation) noise
##' @param rmean name of mean for \code{rfun}
##' @param rpars additional arguments for \code{rfun}
simfun <- function(pars=c(beta=0.2,gamma=0.1,N=1000,i0=0.01),
                   tmax=100,dt=1,
                   rfun=rnbinom,
                   rmean="mu",
                   rpars=list(size=1)
                   ) {
    tvec <- seq(0,tmax,by=dt)
    ss <- SIR.detsim(tvec,pars)
    noiseArgs <- c(setNames(list(length(ss),ss),c("n",rmean)),
                   rpars)
    count <- do.call(rfun,noiseArgs)
    return(data.frame(tvec,count))
}
```

```{r ssex1}
set.seed(101)
s0 <- simfun(rpars=list(size=10))
```


Various ways of fitting ...
```{r fitex1,cache=TRUE}
## get starting values and trajectory based on them
ss0 <- startfun(auto=TRUE,data=s0)
ss2 <- SIR.detsim(s0$tvec,trans.pars(ss0))

## fit and corresponding trajectory
t1 <- system.time(f1 <- fitsir(s0,start=ss0))
ss3 <- SIR.detsim(s0$tvec,trans.pars(coef(f1)))

## GAM fit: match number of degrees of freedom
## glm() lists df as df+2 
##          == length(coef())+1
##   counting for intercept and residual var(?)
m1 <- glm(count~ns(tvec,df=3),
          family=gaussian(link="log"),data=s0)
```

```{r fitex1_plot,echo=FALSE}
par(las=1,bty="l")
plot(count~tvec,data=s0)
lines(s0$tvec,ss2)  ## incidence/prevalence mismatch?
lines(s0$tvec,ss3,col=2)  ## decent fit anyway
lines(s0$tvec,predict(m1,type="response"),col=4)
```


## Preliminary ensemble results

```{r fitsim1,cache=TRUE}
## takes about 1 minute per sim ...
fitfun <- function(data) {
    t1 <- system.time(f1 <- fitsir(data,start=startfun(auto=TRUE,data=data)))
    m1 <- glm(count~ns(tvec,df=3),family=gaussian(link="log"),data=data)
    res <- c(t=unname(t1["elapsed"]),coef(f1),
      nll.SIR=c(-logLik(f1)),nll.gam=c(-logLik(m1)))
    return(res)
}
system.time(print(fitfun(s0)))
```

```{r loadsims}
simfn <- "stochsim.rda"
if (file.exists(simfn)) {
   load(simfn)
} else {
set.seed(101)
res1 <- raply(20,fitfun(simfun(rpars=list(size=10))),
              .progress="text")
}
```

How well does this work?

```{r simhist,echo=FALSE}
par(las=1,bty="l")
with(res1,hist(nll.SIR-nll.gam,col="gray",breaks=20,
               main=""))
```

## Notes/preliminary conclusions

- these are fits to a reasonably well-behaved (although fairly noisy simulation example)
- with the current "autofit" functionality, `fitsir` mostly works OK, doesn't need Latin hypercube for this example
- in 20 sims, GAM/spline fit does slightly better most (90%) of the time, but not a big difference (<2 log-likelihood units)

## To do

- larger sample
- parameters more typical of DR/music-download data: esp. more samples (will presumably make differences *more* significant/favour GAMs more?)
- try factorial experiment:
    - multiple sims
    - range of sample sizes
    - range of true parameter values (Latin hypercube??)

```{r loadsims2}
simfn2 <- "stochsim2.rda"

fitfun.optim <- function(data) {
    t1 <- system.time(f1 <- fitsir.optim(data,start=startfun(auto=TRUE,data=data)))
    m1 <- glm(count~ns(tvec,df=3),family=gaussian(link="log"),data=data)
    res <- c(t=unname(t1["elapsed"]),coef(f1),
      nll.SIR=c(-logLik(f1)),nll.gam=c(-logLik(m1)))
    return(res)
}


if (file.exists(simfn2)) {
   load(simfn2)
} else {
set.seed(101)
res1 <- raply(20,fitfun(simfun(rpars=list(size=10))),
              .progress="text")
}

```