---
title: "Testing auto-start"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r pkgs, message=FALSE}
library(fitsir)
library(bbmle) ## needed for coef() ...
library(splines) ## for ns()
library(plyr)  ## for raply()
library(dplyr)
library(tibble)
library(tidyr)
library(ggplot2); theme_set(theme_bw())
library(gridExtra)
library(ggstance) ## for horizontal violins
library(reshape2)
library(devtools)
load_all("..")
```

```{r load, echo = FALSE}
load("stochsim_swp.rda")
```

Histograms comparing nll of auto-start and true-start:

```{r hist1, message = FALSE, fig.width = 8, fig.height = 4}
resList <- list(res.fitsir, res.fitsir.optim, res.fitsir.optim.nll)
resName <- c("fitsir", "fitsir.optim", "fitsir.optim.nll")
par(mfrow = c(1,3))
hist1 <- lapply(c(1:3), function(i) with(resList[[i]], hist(fit_nll - truefit_nll, main = resName[i])))
```

It seems like `fitsir.optim` and `fitsir.optim.nll` does get stuck sometimes when we're using auto-start values. Can we look at some of them?

```{r badfits}
(badList <- lapply(c(2:3), function(i) with(resList[[i]], which(fit_nll - truefit_nll > 1))))
```

OK, so they look similar but some of them don't overlap. Here are the ones that are common:

```{r bad_common}
(commonbad <- Reduce(intersect, badList))
```

I got rid of the axis labels because it was getting too messy... They're supposed to have different y axes.

```{r plot_badfits, echo = FALSE, fig.width = 10}
source("stochsim_funs.R")
parnames <- names(startfun())
trueparnames <- paste0("truefit_", parnames)
fitparnames <- paste0("fit_start_", parnames)

badfitList <- badData <- badStart <- list()

for(k in 1:length(commonbad)){
    i <- commonbad[k] 
    p <- lhs_df[i,]
    p2 <- c(beta=unname(p["R0"]*p["gamma"]),
            p[2:3],
            i0=unname(p["I0"]/p["N"]))
    true.pars <- with(as.list(p2),
               c(log.beta=log(beta),log.gamma=log(gamma),
                 log.N=log(N),logit.i=qlogis(i0)))
    d <- simfun(pars=p2,tmax=100,dt=1,rpars=list(size=3),seed=101)
    
    autopars <- startfun(data = d, auto = TRUE)
    t <- d$tvec
    auto.I <- SIR.detsim(t, trans.pars(autopars))
    
    tmplist <- list()
    
    for(j in 1:3){
        data <- resList[[j]]
        fpars <- data[i,trueparnames]
        names(fpars) <- parnames
        truefit.I <- SIR.detsim(t, unlist(trans.pars(fpars)))
        fpars2 <- data[i,fitparnames]
        names(fpars2) <- parnames
        autofit.I <- SIR.detsim(t, unlist(trans.pars(fpars2)))
        tmplist[[resName[j]]] <- data.frame(t, truefit.I, autofit.I)
    }
    
    badData[[k]] <- data.frame(t, points = d$count)
    badStart[[k]] <- data.frame(t, auto.I)
    
    badfitList[[k]] <- tmplist
}

mL.fit <- melt(badfitList, id.vars = "t")
mL.data <- melt(badData, id.vars = "t")
mL.start <- melt(badStart, id.vars = "t")

zero_margin <- theme(panel.margin=grid::unit(0,"lines"))
zero_x_margin <-
    theme(panel.margin.x=grid::unit(0, "lines"))

no_axis <- theme(
    axis.ticks.x=element_blank(),
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.y=element_blank(),
    axis.title.y=element_blank(),
    axis.text.y=element_blank()
)

ggplot(mL.fit) + geom_line(aes(t, value, col = L2, linetype = variable, group = interaction(L2,variable)), size = 1) +
    geom_line(data = mL.start, aes(t, value), size = 1.1) +
    geom_point(data = mL.data, aes(t, value), size = 0.6, alpha = 0.4) + 
    facet_wrap(~L1, nrow = 2, scale = "free") +
    no_axis +
    zero_margin +
    zero_x_margin
```

Something else I want to try:

```{r}
which(res.fitsir[,"fit_nll"] - res.fitsir.optim[,"fit_nll"] < -1)
which(res.fitsir[,"fit_nll"] - res.fitsir.optim.nll[,"fit_nll"] < -1)
```

This looks similar to `badList`
